---
title: "Module 6 Lab 5:PCA"
author: "DeBoris Leonard"
date: "2/19/2022"
output: html_document
---
<div style="margin-bottom:50px;">
```{r, include=FALSE}

library(factoextra)
library(FactoMineR)
library(corrplot)
library(RColorBrewer)
library(readxl)
library(ggplot2)


# import sample data
Samples<- read_xlsx("~/Desktop/Ball State/DSCI605/Samples-1.xlsx")
Samples_1 <-Samples[1:6,2:8]
```

The data for this report came from the "Samples-1" excel file. the version 
downloaded contained 6 observations (Rows) and 8 variables (Columns). The 
first column was a character type with the header "Sites". Unable to convert this column to row names, this column was removed in the data wrangling process.

</div>
**PCA on Sample Data**
```{r, echo=FALSE}
#Run PCA on Sample data
Samp_pca <-PCA(Samples_1, scale.unit = FALSE, graph = FALSE)
print(Samp_pca)
```


**Calculating the Eiganvalue**
```{r, echo=FALSE}
#get Eiganvalue of PCA
eig_val <-get_eigenvalue(Samp_pca)
eig_val
```

```{r,fig.cap= "***Figure 1:*** Using fviz_eig Function",fig.align ='center', echo=FALSE}


#visualize data using fviz_eig functions
fviz_eig(Samp_pca, addlabels = TRUE, ylim=c(0,100))
```

Figure 1 shows the variance of the 5 dimensions in the PCA performed on the  
Samples-1 data. We see the most variability in Dimensions 1 (65.7%) and    
Dimensions 2(25.1%). This means we have the majority of our our information 
needed is captured in Principle Components 1 and 2 or "Dimensions 1 and 2".


**Creating "Variable" and "Individual" PCA objects**
```{r, echo=FALSE}
#create variable and ind individual PCA objects
var <- get_pca_var(Samp_pca)
var

ind<-get_pca_ind(Samp_pca)
ind
```


```{r, echo=FALSE,fig.align ='center', fig.cap= '***Figure 2:*** Correlation Plot using corrplot Function'}
#Correlation plot
corrplot(var$cos2, is.corr = FALSE)
```

Our Correlation Matrix indicates that Species E, B and A (in descending order), 
have strong correlations in Dimension 1. We also see that Species C has a 
noteworthy correlation in Dimension 2. The remainders do not show significant 
correlation. 


**Variable PCA**
```{r, echo=FALSE}
#Coordinates 
head(var$coord)
#Cos2: quality on the factor map
head(var$cos2)
#Contributions to the principal components
head(var$contrib)
# Correlation 
head(var$cor)

```


```{r, echo=FALSE,fig.align ='center', fig.cap="***Figure 3:*** Using fviz_pca_var Function"}
#plot the variable PCA
fviz_pca_var(Samp_pca,
             col.var = "contrib", 
             gradient.cols = 'Dark2',
             repel = TRUE     # Avoid text overlapping
             )

```

Figure 3 shows our correlation of variables. We see that Species B,E,F and G  
are the farthest away from the center and have significance. 


**Individual PCA**
```{r, echo=FALSE}
#Coordinates 
head(ind$coord)
#Cos2: quality on the factor map
head(ind$cos2)
#Contributions to the principal components
head(ind$contrib)
```


```{r, echo=FALSE,fig.align ='center', fig.cap= "***Figure 4:*** Using fviz_pca_indFunction"}
#plot the variable PCA
fviz_pca_ind(Samp_pca, col.ind= 'contrib', 
             gradient.cols=c("grey", "#2E9FDF", "#FC4E07" ), repel = TRUE)
```

Figure 4 shows that Sites 2 & 3 are significant variables.



# ***APENDIX***

 
```{r, echo=FALSE,fig.align ='center', fig.cap="***Variable Box chart***"}
#bar chart of correlation
fviz_cos2(Samp_pca, choice = 'var', axes = 1:2)
```



```{r, echo=FALSE,fig.align ='center', fig.cap="***Individual Box chart***"}
fviz_cos2(Samp_pca, choice = 'ind', axes = 1:2)
```

